Popularity, Mixed Matchings, and Self-duality

### 2017-03-27 CO750-ApproximationAlgorithmsTopics/assignments/presentation.tex
proofread slides

The max utility popular matching problem $\text{\alert{(MaxPM)}}$ asks to find a popular matching $\alert{M}$ on $\alert{G}$ which maximizes
$$\sum_{e\in M} w(e)$$
for some weights $\alert{ w: E(G) \rightarrow \Q_+}$.
\end{frame}

\begin{frame}
\frametitle{Fractional Matchings}
We can extend this to the fractional matching case:
$$FM_{G} = \{x \in \R^m_+ : x(\delta(u)) \leq 1, \forall u \in V(G)\}.$$
We extend the voting function by defining for all $\alert{x \in FM_G}$, for all $\alert{u,v' \in V(G)}$:
$$vote_u(x,v) = \sum_{v\in V(G)} x_{uv} vote_u(v,v').$$ 
We extend the vote aggregation defining for all $\alert{x, y \in FM_{G}}$:
$$\Delta(x,y) = \sum_{u \in V(G)} \sum_{uv, uv' \in \delta(u)} x_{uv}y_{uv'} vote_u(v,v').$$
The polytope of all fractional popular matchings is 
$$P_G := \{x \in FM_{G}: \Delta(x,y) \geq 0, \forall y \in FM_G\}.$$
\end{frame}

\begin{frame}
\frametitle{Mixed Popular Matching Example}
In bipartite graphs, each $\alert{x \in FM_G}$ is a convex combination of matchings. But observe that each $\alert{x\in P_G}$ need not be a convex combination of popular matchings.
\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
\begin{figure}
\begin{tikzpicture} 
\node[] (a0) at (0,5) {$a_0$};
    \node[] (a1) at (0,4) {$a_1$};
    \node[] (a2) at (0,3) {$a_2$};

@@

    \path[dashed] (a0) edge (b1);
    \path[dashed] (a0) edge (b2);
    \path[-] (a1) edge (b1);
    \path[-] (a2) edge (b2);
    \path[-,red] (a1) edge (b2);
    \path[-,red] (a2) edge (b1);
\end{tikzpicture}
\end{figure}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
$$ $$
Say each $\alert{a_i}$ has preferences $\alert{b_1 > b_2}$, and each $\alert{b_i}$ has preferences $\alert{a_1 > a_2 > a_0}$.
\end{column}%
\end{columns}
Solid black edges $\alert{S}$ is the only popular matching. If we let $\alert{M}$ be the solid red edges then $$\Pi= \{(S,\frac{1}{2}), (M,\frac{1}{2})\}$$ is a fractional popular matching.
\end{frame}

@@

\begin{frame}
\frametitle{Extended Formulation}
Let $\alert{P'_G}$ be the optimal solutions to $\alert{(LP)}$. Then $\alert{P'_G}$ is described by the set of $\alert{(x,\alpha)}$ such that:
\begin{align*}
\sum_{u \in V(G)} \alpha_u &= 0 \\
\text{s.t.} \alpha_a + \alpha_b &\geq vote_a(b,x) + vote_b(a,x) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u),x) &\forall u \in V(G) \\
\sum_{e \in \delta(u)} x_e &=1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E.
\end{align*}
\end{frame}

@@

So the hypothesis says $\text{\alert{popular matching}}$ is perfect and, in the fractional setting, for each $\alert{u\in A \cup B}$, $\alert{x_{u\ell(u)} = 0}$.

@@

For any $\alert{t \in [0,1)}$ form $\alert{M_t\subseteq E}$ as follows:
\begin{enumerate}
\item Draw a vertical line at distance $\alert{t}$ from left end of $\alert{T}$.
\item For each $\alert{u \in V(G)}$ define $\alert{c_u(t)}$ to be the cell in $\alert{X'_u}$ which the line intersects (or touches the left boundary of).
\item Set $\alert{M_t = \{ uv \in E(G) : u \in V(G) \text{and} v \in c_u(t)\}}$.
\end{enumerate}

@@

Let $\alert{(x, \alpha) \in P'_G}$. Define $\alert{f: P_G \rightarrow P_H}$ by
$$f(x) = z \text{ where } z_{uv'}=z_{u'v} =x_{uv} \text{ and } z_{uu'} = x_{u\ell(u)}.$$
With vector $\alert{\beta}$ such that $\alert{\beta_u = \beta_{u'} = \alpha_u}$ it is easy to check $\alert{Im(f) \subseteq P_H}$.
Define $\alert{h: P_H \rightarrow P_G}$ by 
$$h(z) = y \text{ where } y_{uv} = (z_{uv'} + z_{u'v})/2 \text{ and } y_{u\ell(u)} = z_{uu'}.$$
With witness $\alert{\alpha_u = (\beta_u + \beta_u')/2}$ it is easy to check $\alert{Im(h) \subseteq P_G}$.
$$x = h(f(x)) = \sum_{i=0}^{k-1} h(\lambda_i \chi(M^i)) = \sum_{i=0}^{k-1} \lambda_i h(\chi(M^i)).$$
Each $\alert{h(\chi(M^i))}$ is $\frac{1}{2}$-integral. $\blacksquare$
\end{frame}
\section{Conclusions}

>>>>

The max utility popular matching problem $\text{\alert{(MaxPM)}}$ asks to find a $\text{\alert{popular}}$ matching $\alert{M}$ on $\alert{G}$ which maximizes
$$\sum_{e\in M} w(e)$$
for some weights $\alert{ w: E(G) \rightarrow \Q_+}$.
\end{frame}

\begin{frame}
\frametitle{Fractional Matchings}
We can extend this to the $\text{\alert{fractional matching}}$ case:
$$FM_{G} = \{x \in \R^m_+ : x(\delta(u)) \leq 1, \forall u \in V(G)\}.$$
We extend the voting function by defining for all $\alert{x \in FM_G}$, for all $\alert{u,v' \in V(G)}$:
$$vote_u(x,v) = \sum_{v\in V(G)} x_{uv} vote_u(v,v').$$ 
We extend the vote aggregation defining for all $\alert{x, y \in FM_{G}}$:
$$\Delta(x,y) = \sum_{u \in V(G)} \sum_{uv, uv' \in \delta(u)} x_{uv}y_{uv'} vote_u(v,v').$$
The polytope of all $\text{\alert{fractional popular matchings}}$ is 
$$P_G := \{x \in FM_{G}: \Delta(x,y) \geq 0, \forall y \in FM_G\}.$$
\end{frame}

\begin{frame}
\frametitle{Mixed Popular Matching Example}
In $\text{\alert{bipartite}}$ graphs, each $\alert{x \in FM_G}$ is a convex combination of matchings. But observe that each $\alert{x\in P_G}$ need not be a convex combination of $\text{\alert{popular}}$ matchings.
\begin{columns}[T] % align columns
\begin{column}{.48\textwidth}
\begin{figure}
\begin{tikzpicture} 
\node[] (a0) at (0,5) {$a_0$};
    \node[] (a1) at (0,4) {$a_1$};
    \node[] (a2) at (0,3) {$a_2$};

@@

    \path[dashed] (a0) edge (b1);
    \path[dashed] (a0) edge (b2);
    \path[-] (a1) edge (b1);
    \path[-] (a2) edge (b2);
    \path[-,red] (a1) edge (b2);
    \path[-,red] (a2) edge (b1);
\end{tikzpicture}
\end{figure}
\end{column}%
\hfill%
\begin{column}{.48\textwidth}
$$ $$
Say each $\alert{a_i}$ has preferences $\alert{b_1 > b_2}$, and each $\alert{b_i}$ has preferences $\alert{a_1 > a_2 > a_0}$.
\end{column}%
\end{columns}
Solid black edges $\alert{S}$ is the only $\text{\alert{popular}}$ matching. If we let $\alert{M}$ be the solid red edges then $$\Pi= \{(S,\frac{1}{2}), (M,\frac{1}{2})\}$$ is a $\text{\alert{fractional popular matching}}$.
\end{frame}

@@

\begin{frame}
\frametitle{Extended Formulation}
Let $\alert{P'_G}$ be the optimal solutions to $\alert{(LP)}$. Then $\alert{P'_G}$ is described by the set of $\alert{(x,\alpha)}$ such that:
\begin{align*}
\sum_{u \in V(G)} \alpha_u &= 0 \\
 \alpha_a + \alpha_b &\geq vote_a(b,x) + vote_b(a,x) &\forall ab \in E(G) \\
\alpha_u &\geq vote_u(\ell(u),x) &\forall u \in V(G) \\
\sum_{e \in \delta(u)} x_e &=1 &\forall u \in V(G) \\
x_e &\geq 0 &\forall e \in E.
\end{align*}
\end{frame}

@@

So the hypothesis says each $\text{\alert{popular matching}}$ is perfect and, in the fractional setting, for each $\alert{u\in A \cup B}$, $\alert{x_{u\ell(u)} = 0}$.

@@

For any $\alert{t \in [0,1)}$ form $\alert{M_t\subseteq E}$ as follows:
\begin{enumerate}
\item Draw a vertical line at distance $\alert{t}$ from left end of $\alert{T}$.
\item For each $\alert{u \in V(G)}$ define $\alert{c_u(t)}$ to be the cell in $\alert{X'_u}$ which the line intersects (or touches the left boundary of).
\item Set $\alert{M_t = \{ uv \in E(G) : u \in V(G) \text{ and } v \in c_u(t)\}}$.
\end{enumerate}

@@

Let $\alert{(x, \alpha) \in P'_G}$. Define $\alert{f: P_G \rightarrow P_H}$ by
$$f(x) = z \text{ where } z_{uv'}=z_{u'v} =x_{uv} \text{ and } z_{uu'} = x_{u\ell(u)}.$$
Verified by witness $\alert{\beta}$ such that $\alert{\beta_u = \beta_{u'} = \alpha_u}$.
Define $\alert{h: P_H \rightarrow P_G}$ by 
$$h(z) = y \text{ where } y_{uv} = (z_{uv'} + z_{u'v})/2 \text{ and } y_{u\ell(u)} = z_{uu'}.$$
Verified by witness $\alert{\alpha_u = (\beta_u + \beta_u')/2}$.
$$x = h(f(x)) = \sum_{i=0}^{k-1} h(\lambda_i \chi(M^i)) = \sum_{i=0}^{k-1} \lambda_i h(\chi(M^i)).$$
Each $\alert{h(\chi(M^i))}$ is $\frac{1}{2}$-integral. $\blacksquare$
\end{frame}
\section{Conclusions}
